The reference model approach in feature selection problems \vspace{.1in}

### 2019-05-29 paper/ref_approach.tex
added computational burden subsection; other minor changes

\cite{paper:bodyfat} report the results using the stepwise backward elimination with a significance level of 0.157 with AIC selection fixing abdomen and height to be always included in the model. We implement the selection via projection: the reference model includes all the covariates using as a prior the regularised horseshoe, submodels are explored using a forward search, the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. In order to automatise the procedure, we select the submodel size as the smallest which has an elpd score higher than the reference model with probability 0.05. Figure \ref{fig:inclusion_frequencies} shows the bootstrap inclusions frequencies combining the results from \cite{paper:bodyfat} for the stepwise backward elimination and our experiments for the projection. The projection predictive approach has only two variables with inclusion frequencies above 50 (abdomen is the only included always), the third most included is `wrist' at 49.5 and the fourth is `height' at 27.8, while the stepwise regression has seven features above 50, of which abdomen and height are fixed. Such a lower stability of the stepwise regression can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. The first five selected models have a cumulative frequency of almost 70\% with projpred, whereas with the stepwise regression only of 11.8\%. In addition to that we note that the size of the selected model is much smaller when using projpred.

@@

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|r|r|r|r}
  \hline
 & rmse.full & rmse.sel & size.sel & noisy.sel \\ 
  \hline
projpred & 4.4 & 4.5 & 2 & -  \\
steplm & 4.4 & 4.5 & 7 & - \\
\hline
\hline
projpred & $^{*}$4.4 & $^{*}$4.5 & 2 & 0  \\
steplm & 5.3 & 5.5 & 38 & 32 \\
   \hline
\end{tabular}
\caption{Model predictive performances with original data (top table) and adding noisy features (bottom table). Results with an asterisk using 10-fold CV, otherwise 20-fold CV.}
\label{tab:model_performances}
\end{table}

@@

First we use simulated data in order to vary the difficulty of the selection problem. We rely on the scheme \eqref{eq:simulated_data} using different levels of correlation ($\rho$) and number observation ($n$), precisely $p$ and $k$ are fixed respectively at $1000$ and $100$, whereas $n\in\{50,70,100\}$ and $\rho\in\{0.3,0.5\}$. Lower the correlation level and the number of observation are, more challenging the selection is. As mentioned, this data generation mechanism was already proposed by \cite{paper:projpred}, who also devise a reference model sufficiently good in predictions. It consists in a linear regression using the first five supervised principal components (SPCs) \citep{paper:original_spc, paper:spc} as follows:
\
\begin{equation}
\label{eq:ref_mod}
\begin{aligned}
    Y_{i}|&\boldsymbol{\beta},\sigma^{2},\boldsymbol{u}_{i} \overset{ind}{\sim} N(\boldsymbol{u}_{i}^{T}\boldsymbol{\beta},\sigma^{2}) \quad &i=1,..,n \\
    &\beta_{j}|\!\begin{aligned}[t] &\tau \overset{iid}{\sim} N(0,\tau^{2})\\
    &\tau \sim t_{4}^{+}(0,s_{max}^{-2}) 
    \end{aligned} &j=1,..,5 \\ 
    &\sigma \sim t_{3}^{+}(0,10) \\
\end{aligned}
\end{equation}
where $\boldsymbol{u}_{i}$ stays for the $i$-th SPC and $s_{max}$ denotes the sample standard deviation of the largest SPC. All model are implemented in Stan \citep{paper:stan}. The normal means problem formulation is obtained as explained in Section \ref{comparison} by Fisher-transformation. We label as ``ref'' the reference approach, which consists in using the reference model on top of the computation of the z-values, and ``data'' the standard approach based on the observed data only. In Section \ref{shrinkage-signal} we analyse in the context of full-Bayes methods the different shrinkage and signal estimation provided by the use of a reference model or not. Section \ref{complete-selection} shows results for the complete selection procedures.

@@

We conclude our experiments using the real world dataset body fat, already described in the comparison between the projection predictive approach and the stepwise regression in Section \ref{projection}. Consistently with the previous comparisons, we design the variable selection by normal means problem. We add to the original data noisy uncorrelated covariates up to a total of 1000 features. These covariates are normally distributed and scaled as the original ones. We compute correlations between each variable and the target variable, that is the amount of fat, and transform them by Fisher transformation. The original assumption in order that \eqref{eq:fisher_transformation} holds is that the variables are jointly normally distributed. In our experience the normal approximation is still reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the variance to be one. The methods we compare are those used in Section \ref{complete-selection}, which are the control of the local false discovery rate (loc.fdr), the empirical Bayes median (EB.med) and the selection by posterior credible intervals at level 0.9 (ci.90). The maximum likelihood estimator is used to estimate the variance of the null hypothesis in the local false discovery rate \cite[see][Chap. 6]{book:efron}, whereas the median absolute deviation from zero for the empirical Bayes median method. The selection via posterior credible intervals is done using a regularised horseshoe prior on the parameters for the means, while a standard log-normal  prior for the variance. In order to vary the difficulty of the selection, we bootstrap subsamples of different sizes, going from $n=50$, $70$, $100$ up to $n=251$. Each time we average results over 100 bootstrap subsamples.

>>>>

\cite{paper:bodyfat} report the results using the stepwise backward elimination with a significance level of 0.157 with AIC selection, fixing abdomen and height to be always included in the model. We implement the selection via projection: the reference model includes all the covariates using as a prior the regularised horseshoe, submodels are explored using a forward search and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. In order to automatise the procedure, we select the submodel size as the smallest one which has an elpd score higher than the reference model score with probability 0.05. Figure \ref{fig:inclusion_frequencies} shows the bootstrap inclusions frequencies combining the results from \cite{paper:bodyfat} for the stepwise backward elimination and our experiments for the projection. The projection predictive approach has only two variables with inclusion frequencies above 50\% (abdomen is the only included always), the third most included is `wrist' at 49.5\% and the fourth is `height' at 27.8, while the stepwise regression has seven features above 50\%, of which abdomen and height are fixed. Such a lower stability of the stepwise regression can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. The first five selected models have a cumulative frequency of almost 70\% with projpred, whereas with the stepwise regression only of 11.8\%. In addition to that we note that the size of the selected model is much smaller when using projpred.

@@

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|r|r|r|r}
  \hline
 & rmse.full & rmse.sel & size.sel & noisy.sel \\ 
  \hline
projpred & 4.4 & 4.5 & 2 & -  \\
steplm & 4.4 & 4.5 & 7 & - \\
\hline
\hline
projpred & $^{*}$4.4 & $^{*}$4.5 & 2 & 0  \\
steplm & 5.3 & 5.5 & 38 & 32 \\
   \hline
\end{tabular}
\caption{Model predictive performances with original data (first two rows) and adding noisy features (last two rows). Results with an asterisk using 10-fold CV, otherwise 20-fold CV.}
\label{tab:model_performances}
\end{table}

\hypertarget{computational-burden}{%
\subsection{Computational burden} \label{computational-burden}}

\begin{itemize}
\item add more quantitative results: e.g. times for projpred and steplm
\end{itemize}

We can consider the reference model approach as a family of methods, thus providing an answer to the question what the computational burden is, is not possible, because not unique. It is possible to identify a common and inevitable overhead though, that is the cost of fitting the reference model. In the simple reference model approach we have described at the end of Section \ref{reference-model-approach}, it is actually the only added cost. In general, a full Bayesian model is advisable to have the best predictive performance, however including all the available features can be often quite computationally demanding. In such a case a first possible speed-up could be to use screening or dimensionality reduction techniques, as for example the supervised principal components \citep{paper:original_spc,paper:piironen_spc}. However, in some cases this does not always result in an appreciable speed-up, as when there are many predictive and not correlated features and thus many of them have to be included in the reference model anyway. Other alternatives could be fixing some parameters with their maximum marginal likelihood value (typically referred as empirical Bayes) or even use frequentist approach, if predictions are good enough.

The comparison example between projpred and steplm shows that such additional computational burden can be worth it. In the case of the projection predictive approach it results in sparser submodels, which can result in saving time and general costs using the selected model for future predictions. \ref{tab:model_performances} shows that there could be even an improvement in the prediction ability, thus justifying more additional computational costs in the selection. More benefits obtainable through the reference model are provided in the experiments of Section \ref{comparison}.

@@

First we use simulated data in order to vary the difficulty of the selection problem. We rely on the scheme \eqref{eq:simulated_data} using different levels of correlation ($\rho$) and number observation ($n$), precisely $p$ and $k$ are fixed respectively at $1000$ and $100$, whereas $n\in\{50,70,100\}$ and $\rho\in\{0.3,0.5\}$. Lower the correlation level and the number of observation are, more challenging the selection is. As mentioned, this data generation mechanism was already proposed by \cite{paper:projpred}, who also devise a reference model sufficiently good in predictions. It consists in a linear regression using the first five supervised principal components (SPCs) \citep{paper:original_spc, paper:piironen_spc} as follows:
\
\begin{equation}
\label{eq:ref_mod}
\begin{aligned}
    Y_{i}|&\boldsymbol{\beta},\sigma^{2},\boldsymbol{u}_{i} \overset{ind}{\sim} N(\boldsymbol{u}_{i}^{T}\boldsymbol{\beta},\sigma^{2}) \quad &i=1,..,n \\
    &\beta_{j}|\!\begin{aligned}[t] &\tau \overset{iid}{\sim} N(0,\tau^{2})\\
    &\tau \sim t_{4}^{+}(0,s_{max}^{-2}) 
    \end{aligned} &j=1,..,5 \\ 
    &\sigma \sim t_{3}^{+}(0,10) \\
\end{aligned}
\end{equation}
where $\boldsymbol{u}_{i}$ stays for the $i$-th SPC and $s_{max}$ denotes the sample standard deviation of the largest SPC. All model are implemented in Stan \citep{paper:stan}. The normal means problem formulation is obtained as explained in Section \ref{comparison} by Fisher-transformation. We label as ``ref'' the reference approach, which consists in using the reference model on top of the computation of the z-values, and ``data'' the standard approach based on the observed data only. In Section \ref{shrinkage-signal} we analyse in the context of full-Bayes methods the different shrinkage and signal estimation provided by the use of a reference model or not. Section \ref{complete-selection} shows results for the complete selection procedures.

@@

We conclude our experiments using the real world dataset body fat, already described in the comparison between the projection predictive approach and the stepwise regression in Section \ref{projection}. Consistently with the previous comparisons, we design the variable selection by normal means problem. We add to the original data noisy uncorrelated covariates up to a total of 1000 features. These covariates are normally distributed and scaled as the original ones. We compute correlations between each variable and the target variable, that is the amount of fat, and transform them by Fisher transformation. The original assumption in order that \eqref{eq:fisher_transformation} holds is that the variables are jointly normally distributed. In our experience the normal approximation is still reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the variance to be one. The methods we compare are those used in Section \ref{complete-selection}, which are the control of the local false discovery rate (loc.fdr), the empirical Bayes median (EB.med) and the selection by posterior credible intervals at level 0.9 (ci.90). The maximum likelihood estimator is used to estimate the variance of the null hypothesis in the local false discovery rate \cite[see][Chap. 6]{book:efron}, whereas the median absolute deviation from zero for the empirical Bayes median method. The selection via posterior credible intervals is done using a regularised horseshoe prior on the parameters for the means, while a standard log-normal  prior for the variance. In order to vary the difficulty of the selection, we bootstrap subsamples of different sizes, going from $n=50$ up to $n=251$. Each time results are averaged over 100 bootstrap subsamples.
